Water discharge has a variable impact on sediment transport capacity in subglacial systems

### 2023-08-11 Delaneyetal2023b.tex
updates

\documentclass[11pt]{article}
\usepackage [colon,square]{natbib}
\usepackage {comment}
\usepackage{cite}
\usepackage[mathlines]{lineno} %% <- mathlines turns on line numbering in equations
\usepackage{amsmath} 
\usepackage{printlen}
\setlength{\parindent}{10pt}
\setlength{\parskip}{5pt}
\usepackage[explicit]{titlesec}
\usepackage{helvet}
\usepackage{array}
\usepackage{printlen}
\usepackage[a4paper, margin=1.25cm]{geometry}
\usepackage{color,hyperref}
\catcode`\_=11\relax
\newcommand\email[1]{\_email #1\q_nil}
\def\_email#1@#2\q_nil{%
  \href{mailto:#1@#2}{{\emailfont #1\emailampersat #2}}
}
\newcommand\emailfont{\sffamily}
\newcommand\emailampersat{{\color{red}\small@}}
\catcode`\_=8\relax

@@

ianarburua.delaney@unil.ch}, Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}, Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland}

\end{center}
\paragraph{Keypoints}
\begin{itemize}
\item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, given their subglacial channel's slow evolution.

@@

\item The channels' divergence in transport response may impact sediment dynamics at glacier margins and the interpretation of sediment records.
\end{itemize}

\TODO{Study site description}

\TODO{Consider removing ``case'' ``system'', around \alpine and \icesheet to reduce words}

@@

%\ian{last sentence is a bit rough}

@@

Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
In some of these regions, increased water discharge and glacier melt have been interpreted to be greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
Observed changes to sediment transport in glacierized catchments call for examining the processes controlling sediment discharge in these catchments. 
Yet, the relationship between water discharge and sediment transport in subglacial environments remains minimally discussed and not compared with the subaerial river systems that these glaciers feed.

\TODO{Add Riihimaki and Swift}

@@

In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that the water could carry. Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment. The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that 
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau_t}
    \tau_t \propto w_c\, v^2,
  \end{equation}
\end{linenomath*}
% 
where $w_c$ is the width of the channel.
Following the conservation of mass, the velocity of the water flowing through a  channel is given as 
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q_w}{S},
  \end{equation}
\end{linenomath*}
% 
where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.
\TODO{comment on the squared?}

@@

Yet, the response of water velocity to changing water discharge in subglacial channels is different from subaerial ones.
The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
As a result, in subglacial channels, the water is pressurized by the ice above \citep{rothlisberger1972}.
As a result, the subglacial channel size only evolves over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ changes on a much shorter timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial channel where channel width $w_c$ evolves as well. 
\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig1.pdf}
    \caption{Sketch for the different response that subglacial and subaerial channels have to an increase in water discharge.  Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length. $S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels. The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths evolve $w_{cf}$. Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed. 
      Note that in the subaerial channel parameterization, channels have rectangular shapes (Section~\ref{sect:fluv}).} 
    \label{fig:cartoon}
  \end{figure}
\end{center}
\ian{Come back to Daniel's comments}

As a result, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
%However, some modelss suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width, and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \citep{alley1997,swift2021}.
The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial compared to subaerial systems \citep[e.g.][]{muller1968,richards2003,ganti2016}.
Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels, and implications for sediment discharge records, has been minimally discussed.
Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.

This manuscript aims to evaluate relationship between sediment transport capacity and water discharge in subglacial channels, compared to subaerial ones.
Numerical parameterizations of subglacial and subaerial hydraulics are untilized, in the absence of observations of continuous water velocity measurements and the evolution of subglacial and subaerial channels.
Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
The parameterizations' lumped and simplistic nature is intended to show the disparity between water discharge and sediment transport in subglacial systems  without considering effects such as the upstream drainage network and the role of sediment supply.
Outputs are used to explore differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels. 
With these parameterizations, the manuscript examines the impact of water discharge variability on sediment transport capacity in subglacial and subaerial channels.
The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for sediment transport processes in subglacial channels, compared to subaerial ones, and for interpreting records of sediment transport from glacierized catchments.

@@

\begin{table}[ht]
  \centering
  \caption{Variables, parameters, and constants used in this work. Where two values are given, the first refers to an \alpine{} case (Fieschergletscher) and the second to a glacier marginal to the \icesheet{}case (Leverett Glacier) }
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\ \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
    Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
    Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\

@@

    Ice flow constant &$A$& $5.3\times10^{-24}$ &\unit{Pa}$^{-n}$\,$s^{-1}$\\
    Ice flow exponent &$n$& $3$ &$\mathrm{(-)}$\\
    Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
    Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
    Gradient of channel bed (subaerial) &$\frac{\partial z_c}{\partial x}$ &$0.05$& $\mathrm{(-)}$\\
    Subaerial channel factor & $k$ &$3$ & $\mathrm{s\,m^{-2}}$\\
    Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
    \hline
  \end{tabular}
  \label{table:vpm}

@@

Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003}.
The formulation here does not consider the englacial storage of water.
Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the evolution of the channel size. 
The evolution of subglacial channel size $S_g$ is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:dS_dt}
    \frac{\partial S_g}{\partial t} = C_2 \frac{Q \Delta h}{l} - C_2 (h_{o}-\frac{\Delta h}{2})^n\,S_g,
  \end{equation}
\end{linenomath*}
\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \citep{glen1955}.
The first term on the right side of the equation represents the opening of the channel through frictional heating, while the second term represents the creep closure of the channel from ice deformation.

@@

To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
Here, given  that mass is conserved mass, there is a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, and the Darcy-Weisbach relationship,
the shear stress $\tau_f$ at the river bed is represented as
\begin{linenomath*}
  \begin{equation}
    \label{eq:DW_tau}
    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\frac{\partial z_c}{\partial x}^{\frac{2}{3}},
  \end{equation}
\end{linenomath*}
where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
Channel width $w_{cf}$ is 
\begin{equation}
  \label{eq:wcf}
  w_{cf} = k \, Q_w^\frac{1}{2},
\end{equation}
% 
where $k$ is a constant and the exponent is commonly given as $\frac{1}{2}$ \citep{leopold1953}.
Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
\begin{equation}
  \label{eq:vf}
  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
\end{equation}
% 
As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.

\subsection{Implementation}

The formulation of the glacial channel (Section~\ref{sect:sub_mode}) requires inputs of the ice thickness ($h_{ice}$, Equation\,\ref{eq:dS_dt}), water discharge, $Q_w$,  Hooke angle ($\beta$; Equation~\ref{eq:Hf}), and friction factor (Equations~\ref{eq:dh}~and~\ref{eq:tau}; $f_r$). Values of these parameters are given in Table~\ref{table:vpm}.

In the subaerial channel, the bed slope, $\frac{\partial z_c}{\partial x}$, is needed in Equation~\ref{eq:DW_tau}. The friction factor, $f_f$, and channel shape factor, $k$, are assigned such that they produce reasonable values for water velocity and channel width  (Section~\ref{sect:fluv}).
\TODO{Check what is the ``Reasonable values'' for water velocity. see below also.}

\TODO{Address Daniel's comment: I think that these records need a few more words: where are they measured exactly? How long are the time series? What temporal resolution have they got? Do they cover the entire year, or only the summer months? }

@@

The parameterizations are first applied to a reference test case for each glacier that assumes a semi-circular subglacial channel ($\beta=\pi$) and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and Alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels are presented below.

Second, to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels further, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
Model runs with parameter combinations  are culled if their maximum subglacial water velocity exceeds $2\,$\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity exceeds $1\,$\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some dye tracing measurements \citep[e.g.]{werder2010b,chandler2013}.
Variability in a variable is established from the standard deviation that variable over an aggregation time. 
\ian{Check this again... last sentence needs works (see Daniel's comments)}

\section{Results}
\TODO{needs clarification}

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig2.pdf}
    \caption{Model outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h).". Gray (black) lines in a and e represent the subglacial channel size (water discharge).  In other panels, blue lines represent model outputs from the subglacial channel, while orange lines represent the subaerial channel.
      Plots are shown at $15$\unit{min} intervals.
      Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
      Insets in e-h show peak events at the onset of melt in the \icesheet{} case.
      Note that the x and y axis are different for the \alpine{} and \icesheet{} cases such that they each cover a summer.} 
    \label{fig:model_outs}
  \end{figure}
\end{center}

Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks. Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig3.png}
    \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for the \alpine{}  (a-c) and \icesheet (d-f) cases. Variables have been normalized to mean values for better comparison amongst the model runs.
      Plots are shown at $15$\unit{min} interval.} 
    \label{fig:Qw_vari}
  \end{figure}
\end{center}

Each water discharge value in the subaerial channel is associated with a single water velocity, shear stress and width-integrated shear stress (Figure~\ref{fig:Qw_vari}).
Conversely in the subglacial channel,  water velocity, shear stress and width-integrated shear stress cab vary substantially for a given water discharge (Figure~\ref{fig:Qw_vari}).
For instance, over $15$ minute time periods at \alpine{}, the range of water velocities and shear stress spans from those below the mean to those almost $2$ times the mean velocity.
Very high values of water velocity and shear stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
At \icesheet, reduced diurnal variability in the hydrograph  (Figure~\ref{fig:model_outs} e) results in smaller variations in water velocity and shear stress compared to \alpine{} (Figure~\ref{fig:Qw_vari}).
However in \icesheet, mean values of water velocity can occur at water discharges between roughly $50$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $225$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the observed range of water discharge. 
\TODO{add sentence about why the two cases were different}

>>>>

\documentclass[11pt]{article}
\usepackage [colon,square]{natbib}
\usepackage[mathlines]{lineno} %% <- mathlines turns on line numbering in equations
\usepackage{amsmath} 
\usepackage{printlen}
\setlength{\parindent}{10pt}
\setlength{\parskip}{10pt}
\usepackage[explicit]{titlesec}
\usepackage{helvet}
\usepackage{array}
\usepackage{printlen}
\usepackage[a4paper, margin=1.3cm]{geometry}
\usepackage{color,hyperref}
\catcode`\_=11\relax
\newcommand\email[1]{\_email #1\q_nil}
\def\_email#1@#2\q_nil{%
  \href{mailto:#1@#2}{{\emailfont #1\emailampersat #2}}
}
\newcommand\emailfont{\sffamily}
\newcommand\emailampersat{{\color{red}\small@}}
\catcode`\_=8\relax

@@

    ianarburua.delaney@unil.ch},
  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland},
  Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland},

\end{center}
\paragraph{Keypoints}
\begin{itemize}
\item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size, given their subglacial channel's evolution.

@@

\item The channels' divergence in sediment transport response represents multiple processes represented by water discharge in sediment transport.
\end{itemize}

@@

@@

Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
In some of these regions, increased water discharge and glacier melt have been interpreted to be greater sediment transport capacity \citep{bendixen2017,costa2017,li2021}.
Observed changes to sediment transport in glacierized catchments require examining the processes controlling sediment discharge in these catchments and its variations with water discharge  \citep[e.g.][]{riihimaki2005,swift2005}.

@@

In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that the water can carry.
Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that 
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau_t}
    \tau_t \propto w_c\, v^2,
  \end{equation}
\end{linenomath*}
% 
where $w_c$ is the width of the channel.
Following the conservation of mass, the velocity of the water flowing through a  channel is given as 
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q_w}{S},
  \end{equation}
\end{linenomath*}
% 
where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.

@@

Yet, the response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
As a result, the subglacial channel size only evolves over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much shorter timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial channel where channel width $w_c$ evolves as well. 
\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig1.pdf}
    \caption{Sketch for the different response that subglacial and subaerial channels have to an increase in water discharge.  Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length. $S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels. The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths evolve $w_{cf}$. Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed. 
      Note that in the subaerial channel parameterization, channels have rectangular shapes (Section~\ref{sect:fluv}).} 
    \label{fig:cartoon}
  \end{figure}
\end{center}
% \ian{Come back to Daniel's comments}

As a result, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial systems \citep[e.g.][]{muller1968,richards2003,swift2005,ganti2016}.
Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels, and implications for sediment discharge records, has been minimally discussed.
Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.

This manuscript aims to evaluate relationship between sediment transport capacity and water discharge in subglacial channels, compared to subaerial ones.
Numerical parameterizations of subglacial and subaerial hydraulics are untilized, in the absence of observations of continuous water velocity measurements and the evolution of subglacial and subaerial channels.
Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
The parameterizations' lumped and simplistic nature is intended to show the disparity between water discharge and sediment transport in subglacial systems  without considering effects such as the upstream drainage network and the role of sediment supply.
Outputs are used to explore differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels. 
Furthermore, outputs are compared to sediment discharge data from these two catchments.
The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in records of sediment transport from glacierized catchments.

\section{Study sites and data}
\label{sect:ss_data}

% \TODO{Address Daniel's comment: I think that these records need a few more words: where are they measured exactly? How long are the time series? What temporal resolution have they got? Do they cover the entire year, or only the summer months? }

Water and suspended sediment data are leveraged from two sites in this study.
Water discharge and suspended sediment concentration were collected at a $1$\unit{min} interval from the Fieschergletscher in the Swiss Alps  continuously over the period 2014--2021 ($46^\circ\,29'\,07"$ N, $8^\circ\,08'\,34"$ E).
A description of the data and its collection is available in \citet{felix2022} and \citet{felix2021}.

Additionally, water discharge and suspended sediment concentration were collected roughly $2$\unit{km} downstream from Leverett Glacier in Greenland from over  periods from June to September from 2009 to 2012 ($67^\circ\,03'\,50"$ N, $50^\circ\,12'\,59"$ W).
This data was presented in part in \citet{cowton2012} and is available from \citet{tedstone2017} at a $5$\unit{min} time interval.
Note that in some case the removal and relocation of sensors created variations in the sediment concentration that do not result from natural processes.

@@

\begin{table}[ht]
  \centering
  \caption{Variables, parameters, and constants used in this work. Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier }
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\ \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
    Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
    Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\

@@

    Ice flow constant &$A$& $5.3\times10^{-24}$ &\unit{Pa}$^{-n}$\,$s^{-1}$\\
    Ice flow exponent &$n$& $3$ &$\mathrm{(-)}$\\
    Friction factor (glacier) & $f_r$ & $4$ & $\mathrm{(-)}$ \\
    Friction factor (subaerial) & $f_p$ & $3$ & $\mathrm{(-)}$\\
    Gradient of channel bed (subaerial) &$\nabla z_c$ &$0.05$& $\mathrm{(-)}$\\
    Subaerial channel factor & $k$ &$3$ & $\mathrm{s\,m^{-2}}$\\
    Channel geometry exponent &$e$& $\frac{1}{2}$&$\mathrm{(-)}$ \\
    \hline
  \end{tabular}
  \label{table:vpm}

@@

Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003}.
The formulation here does not consider the englacial storage of water.
Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the evolution of the channel size. 
The evolution of subglacial channel size $S_g$ is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:dS_dt}
    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 (h_{o}-\frac{\Delta h}{2})^n\,S_g,
  \end{equation}
\end{linenomath*}
\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \citep{glen1955}.
The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.

@@

To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
Here, given  that mass is conserved mass, there is a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, and the Darcy-Weisbach relationship,
the shear stress $\tau_f$ at the river bed is represented as
\begin{linenomath*}
  \begin{equation}
    \label{eq:DW_tau}
    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
  \end{equation}
\end{linenomath*}
where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
Channel width $w_{cf}$ is 
\begin{equation}
  \label{eq:wcf}
  w_{cf} = k \, Q_w^\frac{1}{2},
\end{equation}
% 
where $k$ is a constant and the exponent is commonly given as $\frac{1}{2}$ \citep{leopold1953}.
Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
\begin{equation}
  \label{eq:vf}
  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
\end{equation}
% 
As in Section~\ref{sect:sub_mode}, the width-integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.

\subsection{Implementation}
\label{sect:imp}

@@

The parameterizations are first applied to a reference test case for each glacier that assumes  subglacial channel with $\beta=\frac{\pi}{2}$ and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and Alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
In both subaerial channels, the bed slope, $\nabla z_c$ is $0.05$.
Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels from this formulation are presented below.

Second, to characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels further, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
Model runs with parameter combinations  are culled if their mean subglacial water velocity over the season lies outside of the range between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outside of the range $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some measurements \citep[e.g.]{werder2010b,magnusson2012,chandler2013}.
The standard deviation of a model output over the course of the season is used to quantify the variability of the model output.

Lastly, model outputs are compared with record of sediment discharge from the two glaciers using Spearman rank correlation to reduce the impact of non-linear response of sediment discharge to the output.

\section{Results}

\subsection{Role of water discharge in sediment transport capacity}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\linewidth]{Fig2.pdf}
  \caption{Model outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h).". Gray (black) lines in a and e represent the subglacial channel size (water discharge).  In other panels, blue lines represent model outputs from the subglacial channel, while orange lines represent the subaerial channel.
    Data in plots are at $15$\unit{min} intervals.
    Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
    Insets in e-h show peak events at the onset of melt in \icesheet{}.
    Note that the x and y axis are different for \alpine{} and \icesheet{} such that they each cover a summer.} 
  \label{fig:model_outs}
\end{figure}
Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks. Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).

Over the course of the season, large peaks in water velocity and shear stress, along with width-integrated shear stress, occur in response to increases in water discharge in both cases.
In the subglacial channel, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
Conversely, in the subaerial channel, peaks in these variables occur when water discharge is at its highest value.
As a result, peaks in the subglacial sediment transport variables occur  when water discharge change is greatest, as opposed to when the maximum amount of water discharge occurs.
The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge. 
Here, sediment could be deposited in the proglacial area as the water depressurizes and remobilized when sediment transport capacity peaks in the subaerial system.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\linewidth]{Fig3.png}
  \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{}  (a-c) and \icesheet (d-f). Variables have been normalized to mean values for better comparison amongst the model runs.
    $R_g$ ($R_f$) show the Spearman rank correlation values for the subglacial (subaerial) model outputs.
    Plots are shown at $15$\unit{min} interval.} 
  \label{fig:Qw_vari}
\end{figure}

Each water discharge value in the subaerial channel is associated with a single water velocity, shear stress and width-integrated shear stress, resulting in a perfect rank correlation between water discharge and the model output (Figure~\ref{fig:Qw_vari} a--f).
Conversely in the subglacial channel,  water velocity, shear stress and width-integrated shear stress can vary substantially for a given water discharge and width-integrated shear stress varies with respect to the other two model outputs as well (Figure~\ref{fig:Qw_vari} a--f).
For instance, very high values of water velocity and shear stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
In \icesheet, mean values of water velocity can occur at water discharges between roughly $145$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $215$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the much of the water discharge range.
